Title
========================================================

The data for this project comes from a personal activity monitoring device. I first load the data from an online backup and proceed to preprocess it for analysis.I first remove the NA values and reclassifiy the date column as an R date object. I create an extra variable called Weekdays to help make the analysis simple. 
```{r}
if(!file.exists("./data")){dir.create("./data")}
fileurl = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileurl, destfile = "./data/data.zip")
data = unzip("./data/data.zip")
data = read.csv(data, header = T)

data$date = as.Date(data$date)
data$weekdays = as.factor(weekdays(data$date))
mydata = na.omit(data)
```
### Histogram of Steps taken each day
```{r}
sums = data.frame(sums = with(mydata, tapply(steps, date, sum)))
barplot(sums$sums, space = F,
        ylab = "Number of steps per day",
        main = "Histogram of Steps taken each day")
```
The mean and median steps taken each day are listed below.
```{r}
Means = with(mydata, tapply(steps, date, mean))
Medians = with(mydata, tapply(steps, date, median))
```
```{r}
Means
Medians
```


This is a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.
```{r}
Interval = data.frame(Average = with(mydata, tapply(steps, as.factor(interval), mean)))
plot(Interval, type = "l",
     xlab = "5 minute Intervals",
     ylab = "Average number of steps taken",
     main = "Time Series Plot of the average steps taken per interval")
```
There are a total of 2304 cases with missing data.
```{r}
n = nrow(data) - nrow(mydata)
n
```

I use the averages per day to fill in the missing values. So for days where no data was recorded, I put in 0.
```{r}
Means = with(mydata, tapply(steps, date, mean))
for(i in 1:nrow(data)){
    if(is.na(data[i, 1])){
        date = data[i, 2]
        data[i, 1] = Means[date]
    }
    if(is.na(data[i, 1])){
        data[i, 1] = 0
    }
}
```
This new dataset has missing values filled in using averages for that day.
```{r}
head(data, n = 10)
tail(data, n = 10)
```

### Histogram of Steps taken per day of the week
```{r}
sums = data.frame(sums = with(data, tapply(steps, weekdays, sum)))
barplot(sums$sums, space = F,
        ylab = "Number of steps per day")
```
The mean and median steps taken each day are listed below.
```{r}
Means = with(data, tapply(steps, date, mean))
Medians = with(data, tapply(steps, date, median))
```
```{r}
Means
Medians
```
I created a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.I then used this factor variable to create the plot below.
```{r}
data$week = NA
data$weekdays = as.character(data$weekdays)
for(i in 1:nrow(data)){
    if(data[i,4] == "Monday"){data[i,5] = "weekday"}
    if(data[i,4] == "Tuesday"){data[i,5] = "weekday"}
    if(data[i,4] == "Wednesday"){data[i,5] = "weekday"}
    if(data[i,4] == "Thursday"){data[i,5] = "weekday"}
    if(data[i,4] == "Friday"){data[i,5] = "weekday"}
    if(data[i,4] == "Saturday"){data[i,5] = "weekend"}
    if(data[i,4] == "Sunday"){data[i,5] = "weekend"}
}

df1 = subset(data, data$week == "weekend")
Interval1 = data.frame(Average = with(df1, tapply(steps, as.factor(interval), mean)),
                       week = "weekend",
                       interval = names(with(df1, tapply(steps, as.factor(interval), mean))))
df2 = subset(data, data$week == "weekday")
Interval2 = data.frame(Average = with(df2, tapply(steps, as.factor(interval), mean)),
                       week = "weekday",
                       interval = names(with(df1, tapply(steps, as.factor(interval), mean))))
df = rbind(Interval1, Interval2)
library(lattice)
df$week = as.factor(df$week)
xyplot(df$Average ~ df$interval | df$week, data = df,
       type = "l",
       xlab = "Interval",
       ylab = "Number of steps",
       layout=c(1,2))
```

